\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
%%%%%%%%% options for the file macros.tex

\def\showauthornotes{1}
\def\showkeys{0}
\def\showdraftbox{0}
\let\pref=\prettyref

\input{macros}
\allowdisplaybreaks

\title{CSC2240 Class Project -- Oblivious Subspace Embeddings}
\author{Aida Ramezani \& Marc-Etienne Brunet}
\date{December 2022}

\begin{document}
\setlength\parindent{0pt}
\setlength\parskip{0.5em}

\maketitle

\section{Introduction}
Oblivious subspace embeddings (OSEs) have numerous applications in approximation and sketching algorithms.
They are also interesting in their own right,
challenging our intuitions about information compression, 
and the geometry of high dimensional space.

\subsection{Outline}
In this project we cover the following:

\begin{enumerate}
    \item Review of the Johnson-Lindenstrauss (JL) Lemma seen in class. 
    \item Introduction to oblivious subspace embeddings (OSEs).
    \item Application of OSEs to linear regression.
    \item Motivation for sparse OSEs.
    \item Review of the trace inequalities proof from ~\cite{cohen2016nearly}.
\end{enumerate}

\subsection{Notation}
To facilitate readability, we use the following notation throughout. 
\begin{table}[]
    \centering
    \begin{tabular}{c|l}
        $n$ & dimension in original space \\
        $m$ & dimension of embedding \\
        $d$ & dimension of subspace \\
        $k$ & number of vectors preserved by JL \\
        $\Pi$ & Embedding matrix $\in \rea^{m \times n}$
    \end{tabular}
    % \caption{Caption}
    \label{tab:my_label}
\end{table}


\section{Review of the JL-Lemma}
\input{review_jl.tex}

\section{Oblivious Subspace Embeddings}
\input{subspace.tex}

\section{Application to linear regression}
\subsection{Motivation and setup}
\input{regression.tex}

\subsection{OSEs for approximate regression}
\input{regression_proof.tex}

\subsection{Why do we want \textit{sparse} subspace embeddings?}
\input{sparsity.tex}

\subsection{Proof sketch of bounds}
\input{proof.tex}


\bibliographystyle{apalike}
\bibliography{bibs}

\newpage
\section*{Appendix}
\appendix
\section{Question about Woodruff's proof}
\label{app:question}
\input{question.tex}

\end{document}
