\documentclass{article}
\usepackage[utf8]{inputenc}
%%%%%%%%% options for the file macros.tex

\def\showauthornotes{1}
\def\showkeys{0}
\def\showdraftbox{0}
\let\pref=\prettyref


\input{macros}

%%%%%%%%% Authornotes
\newcommand{\Snote}{\Authornote{S}}
\newcommand{\Scomment}{\Authorcomment{S}}
\newcommand{\Sfnote}{\Authorfnote{S}}


\title{CSC2240 Class Project -- Oblivious Subspace Embeddings}
\author{Aida Ramezani \& Marc-Etienne Brunet}
\date{November 2022}

\usepackage[a4paper, total={4in, 8in}]{geometry}
\begin{document}



\maketitle

\section{Introduction}
\subsection{Outline}
\begin{enumerate}
    \item [{[}3 min{]}] What are subspace embeddings? Compare to JL lemma.
    \item [{[}3 min{]}] Why are they interesting? Motivate with regression, compare to JL lemma. 
    \item [{[}3 min{]}] Why do we want sparse subspace embeddings?
    \item [{[}11 min{]}] Sketch the simpler Matrix-Chernoff bounds from Section ``Our Approach''.
    \item [{[}5 min{]}] Demo - either regression or NLP. 
    \item [{[}5 min{]}] Q \& A
\end{enumerate}

\subsection{Notation}
\begin{tabular}{c|l}
    $n$ & dimension in original space \\
    $m$ & dimension of embedding \\
    $\Pi$ & embedding matrix $\in \rea^{m \times n}$ \\
    $k$ & number of vectors preserved by JL \\
    $d$ & dimension of subspace \\
\end{tabular}

\subsection{What are oblivious subspace embeddings (OSEs)?}
\input{subspace.tex}


\subsection{Why are subspace embeddings useful?}
\input{regression.tex}

\subsection{Why do we want \textit{sparse} subspace embeddings?}
At a high level, we are interested in a sparse subspace embeddings,
because the procedure for sampling a dense embedding matrix $L$,
then multiplying it through to embed the problem, 
may offset the computational benefit of embedding the problem in the first place. 
% What do we mean by sparse?
% How can sparsity further help improve time complexity?
\input{sparsity.tex}

\subsection{Sketch of proof of bounds}
\input{proof.tex}

\newpage
\bibliographystyle{apalike}
\bibliography{bibs}
\end{document}
